---
title: "R Markdown summary"
author: "Lynnstacy Kegeshi"
output: 
  pdf_document: default
  word_document: default
  html_document: default
date: "2024-10-25"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Introduction to R Markdown

**R Markdown** is an authoring framework for data science, combining code, narrative, and visualizations in a single, reproducible document. It allows users to seamlessly blend Markdown for formatting with R code chunks, enabling dynamic reports, presentations, and websites.

## Key Functionalities of R Markdown

R Markdown is highly versatile, offering several key functionalities for data scientists and analysts:

- **Dynamic Document Creation**: R Markdown supports integrating code and text, enabling users to write code and instantly view its output.
  
- **Reproducible Analysis**: Since the code and documentation are in one place, R Markdown documents are highly reproducible. They can be rerun at any time to regenerate the document with updated data.

- **Visualizations**: R Markdown can embed plots and visualizations directly in the document, which is essential for data reporting and analysis.

- **Support for Various Output Formats**: Users can knit their documents into various formats, such as HTML, PDF, Word, or even slideshows.

- **Customization**: R Markdown allows for extensive customization using YAML headers and CSS, so users can modify the layout, style, and format of their documents.

## Example of a Plot in R Markdown

Here’s an example of how R Markdown can dynamically generate a plot:

```{r}
# Generate a sample plot
plot(cars, main = "Speed vs Stopping Distance", col = "blue")
```

